Ore Polynomials in Sage

### 2013-06-13 paper/main.tex
add a table of methods and an example for differential operators

@@

@@

In some cases (for example when the base ring is $\mathbb{Z}$ or
$\mathbb{Z}[x]$), isolated values of a sequence can be
computed asymptotically faster for large $n$ than by listing all
values, using the binary splitting technique.
The \verb|forward_matrix_bsplit| method, called with argument~$n$,
returns a matrix $P$ and a denominator $Q$ such that $P / Q$ multiplied a
column vector of initial values $c_0, c_1, \ldots$
yields $c_n, c_{n+1}, \ldots$.
This way, computing $10^5$ digits of $\mathrm{e}$ takes a fraction of a second:

>>>>

\subsection{Differential Operators}

Useful methods available for differential operators:

\begin{center}
  \begin{tabular}{|l|p{.55\hsize}|}
    \hline
    Method name & short description \\\hline
    \verb|to_S()| & converts to a recurrence operator for the Taylor series solutions at the origin \\
    \verb|to_F()| & converts to a difference operator for the Taylor series solutions at the origin \\
    \verb|to_T()| & rewrites in terms of the Euler derivative \\
    \verb|annihilator_of_integral()| & converts an annihilator for $f(x)$ to one for $\int f(x) dx$\\
    \verb|annihilator_of_composition()| & converts an annihilator for $f(x)$ to one for $f(a(x))$ where $a(x)$ is algebraic over the base ring\\
    \verb|power_series_solutions()| & computes power series solutions of this operator\\
    \verb|generalized_series_solutions()| & computes generalized series solutions of this operator \\ \hline
  \end{tabular}
\end{center}

We compute an annihilator for the error function $\frac{2}{\sqrt{\pi}} \int_0^x e^{-t^2} dt$,
starting from the differential equation for $e^x$, and produce the
recurrence for the Taylor series coefficients at the origin:

\begin{sageexample}
  sage: R.<x> = ZZ['x']; A.<Dx> = OreAlgebra(R, 'Dx')                            
  sage: (Dx - 1).annihilator_of_composition(-x**2)
  Dx + 2*x
  sage: (Dx + 2*x).annihilator_of_integral()                                     
  Dx^2 + 2*x*Dx
  sage: (Dx^2 + 2*x*Dx).to_S(OreAlgebra(ZZ['n'], 'Sn'))                          
  (n^2 + 3*n + 2)*Sn^2 + 2*n
  sage: (Dx^2 + 2*x*Dx).power_series_solutions(10)                               
  [x - 1/3*x^3 + 1/10*x^5 - 1/42*x^7 + 1/216*x^9 + O(x^10), 1 + O(x^10)]
\end{sageexample}

@@

Useful methods available for recurrence operators:

@@

In some cases (for example when the base ring is $\mathbb{Z}$ or
$\mathbb{Z}[x]$), isolated values of a sequence can be
computed asymptotically faster for large $n$ than by listing all
values, using the binary splitting technique.
The \verb|forward_matrix_bsplit| method, called with argument~$n$,
returns a matrix $P$ and a denominator $Q$ such that $P / Q$ multiplied a
column vector of initial values $c_0, c_1, \ldots$
yields $c_n, c_{n+1}, \ldots$.
This way, computing $10^5$ digits of $e$ takes a fraction of a second:
